\papertitle

### 2020-01-27 paper/paper.tex
More changing in the text.

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage[colorlinks=true,urlcolor=red]{hyperref}
\usepackage{url}
\usepackage{float}
\usepackage{cleveref}
\usepackage{todonotes}
\usepackage{comment}
\usepackage{ulem}

@@

@@

From the user perspective, the workflow can be described by the following categories:

\item[Input] The information inserted in the workflow. It is composed of datasets describing, for instance:

\item The actual input data, for instance, the dates for the experiments, how many times the workflow has to run, etc.

\item The dependencies among the tasks.

\item The I/O information, i.e., where the datasets will be stored in each step of the workflow.

\item The information about the lifetime of the final outcome.

\item The steps in which manual decisions have to be made.

For simplicity, we will call these datasets as \texttt{input.ds}.

\item[Task 1] The task is specified by the user and the information we have about it is:

\item The data input.

\item The data output.

\item How long the data has to be stored.

\item The dependencies regarding the other tasks.

In the case of Task 1, we have:

\item Task\,1 has no prior dependencies other than the general input.

\item Task\,1 is a dependency for running Task\,2 and Task\,3.

\item For simplification purposes, we will assume that Task\,1 has the datasets \texttt{input1.ds}, \texttt{output1.ds}, and \texttt{storage1.ds} as descriptors.

\item[Task 2]

\item Task\,2 is dependent on the outcomes of Task\,1.

\item Task\,2 has the datasets \texttt{input2.ds}, \texttt{output2.ds}, and \texttt{storage2.ds} as descriptors.

\item[Task 3]

\item Task 3 dependent on the outcomes of Task\,1 and Task\,2. It does not require any other task to be started and, after finished, no other task will use its results other than the overall output.

\item Task 3 has the datasets \texttt{input3.ds}, \texttt{output3.ds}, and \texttt{storage3.ds} as descriptors.

\item[Product]

A product is any of the possible outcomes of a task. Usually is composed of datasets that will be used by other tasks or generate the final output.

\item[Final Output]

The final output consists of all datasets generated by the workflow. Those datasets have to be stored for a specific amount of time previously described in one of the input datasets.

For simplicity, we will call these datasets as \texttt{output-workflow.ds}. In this specific example, it is composed by the datasets \texttt{output1.ds}, \texttt{output2.ds}, and \texttt{output3.ds}.

For this work, the datasets belonging to \texttt{input-workflow.ds} are the datasets inserted as input for the running Cylc. The Cylc tool receives, as predefined, information about the storage and dependencies of the datasets. This information is used to construct and run the complete workflow, but it is not used to optimise intermediate steps.
The configuration is already predefined, and it cannot change. For instance, if it is established that Task\,2 depends on Task\,1, then we have to wait for Task\,1 to be completed to start running Task\,2.
But several options can be optimised if we know the whole workflow in advance. Thinking what about I/O, for instance, no decision is being made about an optimal way to store the data. Consider, as an example, that the datasets from \texttt{output1.ds} are divided into \texttt{output1-final.ds} and \texttt{output1-task2.ds}. Then, it is clear that \texttt{output1-task2.ds} will be required to run Task\,2, and, because of that, should be put in fast storage to quickly being retrieved, or even kept in local storage. The \texttt{output1-final.ds} dataset, however, can go directly to tape, if the lifetime of the final output has to be stored for a long time, but no access is required from the current workflow.

@@

\item Task 1 usually requires ten hours running in a pc with standard configuration\footnote{Intel Core i5, 8 GB RAM, and 500 GB internal storage drive, for instance.}, requires 100 GB of storage that has to be persistent for no more than one week.

\item Task 2 usually requires two hours running in a pc with standard configuration, requires 150 GB of storage that has to be persistent for no more than two weeks week.

\item Task 3 usually requires five hours running in a pc with standard configuration, and it does not require extra storage other than the storage for the final results.

@@

Datasets generated by Task \textbf{A} and used by Task \textbf{B} will have the same physical location in the system because there is one fixed script responsible for setting the configuration. Again, such configuration was done manually and with restricted information about the storage and system architecture. It would be interesting to explore the options for the datasets and have them placed in different storage systems. If the lifetime of such dataset is, for instance, two cycles, alternate the datasets into two scratch dataset systems is something that would be a simple job for a tool working on generating filenames and storage. However, manually, that implies at having at least two scripts with information about the different storage and extra work to have better utilisation of the available resources.

@@

\begin{figure}[H]
  \centering
  \includegraphics[scale=1.4]{stages}
  \caption{Stages}
  \label{fig:work2}
\end{figure}

@@

@@

@@

@@

@@

  Now, Cylc will call into EWIOScheduler to potentially reprioritise the order of execution of tasks.
  The EWIOScheduler will read the suite file and the suite-io.rc file to make suggestions but won't enforce anything.

  Potentially, Cylc will change the metadata of the script telling Slurm to place a job on the same nodes as a previous job to reuse local data.
  Potentially, Cylc may also add as metadata information to migrate data.

  Then runs Slurm to queue the task.

  If migration has to be performed before or after a job is run, it would do so.

  The way filenames are generated now slightly different, using a new command that then uses ESWIOSchedule.
  The guy reads the suite file(s) and generates information for ESDM where to prioritise to place or access data. The filename hence encoded additional information.
  Potentially, we generate an esdm.conf file to reflect the specific hardware we want to use.

  The application may use NetCDF, XIOS, or ESDM to perform the I/O.

  Knowing the potential targets, ESDM loads the system configuration to schedule application-local I/O.
  The application uses ESDM with the script filename, ESDM extracts the long-term schedule information and considers it during "application-local" I/O scheduling.

  \item A data management service may run and purge or migrate data according to user-specifications of the information life cycle.

@@

\begin{figure}[H]
  \centering
  \includegraphics[scale=0.6]{workflow-all}
  \caption{Workflow 2.}
  \label{fig:work2}
\end{figure}

@@

>>>>

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage[colorlinks=true,urlcolor=red]{hyperref}
\usepackage{url}
\usepackage{float}
\usepackage{cleveref}
\usepackage{todonotes}
\usepackage{comment}
\usepackage{ulem}
\usepackage{array}

@@

From the user perspective, a general workflow can be described by three categories described in Table \ref{tab:workflow}.

\begin{table}[H]
\centering
\begin{tabular}{|m{2cm}|m{5.5cm}|m{3.5cm}|m{3.5cm}|}
\hline
& \parbox{5.5cm}{\centering \textbf{Input}} & \parbox{3.5cm}{\centering \textbf{Tasks}} & \parbox{3.5cm}{\centering \textbf{Products}} \\ \hline \hline

Description & Information inserted in the workflow. & The tasks specified by the user. & Possible outcomes of the tasks. \\ \hline

File & \texttt{input.ds} & \texttt{tasks.ds} & \texttt{output.ds} \\ \hline

Composed of & Datasets/Scripts & Datasets/Scripts & Datasets \\ \hline

Information &

\begin{tabular}{m{5cm}}
\textbullet\ Dates for the experiments, how many times the workflow has to run, etc.\\
\textbullet\ Dependencies among the tasks.\\
\textbullet\ Information about the lifetime of the final outcome.\\
\textbullet\ Steps in which manual decisions have to be made.\\
\textbullet\ {{\color{cyan}{I/O information. (Where the datasets will be stored in each step of the workflow.)}}}\\
\end{tabular}

&

\begin{tabular}{m{3cm}}
\textbullet\ The data input.\\ \\
\textbullet\ The data output.\\ \\
\textbullet\ The data lifetime.\\ \\
\textbullet\ The dependencies regarding the other tasks.\\ \\
\textbullet\ {{\color{cyan}{I/O information.}}}
\end{tabular}

&

\begin{tabular}{m{3cm}}
\textbullet\ Datasets that will be used by other tasks.\\ \\
\textbullet\ Datasets that will generate the final output.\\
\end{tabular}

\\ \hline
\end{tabular}
\caption{Components of the workflow in the user perspective}
\label{tab:workflow}
\end{table}

For this work, Cylc is the essential tool regarding not only the execution of the workflow, but the exchange of information with the EWIOScheduler about I/O and storage.
Cylc receives, as predefined, information about the storage and dependencies of the datasets. The current information is used to construct and run the complete workflow, but it is not employed in optimising intermediate steps.
The configuration is already predefined, and it cannot change. For instance, if it is established that Task\,2 depends on Task\,1, then we have to wait for Task\,1 to be completed to start running Task\,2.
Thinking what about I/O, no decision is being made about an optimal way to store the data.
Consider, as an example, that the datasets from \texttt{product1.ds} (the final product of task 1) are divided into \texttt{product1-final.ds} and \texttt{product1-task2.ds}. Then, it is clear that \texttt{product1-task2.ds} will be required to run Task\,2, and, because of that, should be put in fast storage to quickly being retrieved, or even kept in local storage. The \texttt{product1-final.ds} dataset, however, can go directly to tape, if the lifetime of the final output has to be stored for a long time, but no access is required from the current workflow.
The bottom line here is: several options can be optimised if we know the whole workflow in advance.

@@

@@

\item Task\,1 usually requires ten hours running in a pc with standard configuration\footnote{Intel Core i5, 8 GB RAM, and 500 GB internal storage drive, for instance.}, requires 100 GB of storage that has to be persistent for no more than one week.

\item Task\,2 usually requires two hours running in a pc with standard configuration, requires 150 GB of storage that has to be persistent for no more than two weeks week.

\item Task\,3 usually requires five hours running in a pc with standard configuration, and it does not require extra storage other than the storage for the final results.

@@

Datasets generated by Task\,\textbf{A} and used by Task\,\textbf{B} will have the same physical location in the system because there is one fixed script responsible for setting the configuration. Again, such configuration was done manually and with restricted information about the storage and system architecture. It would be interesting to explore the options for the datasets and have them placed in different storage systems. If the lifetime of such dataset is, for instance, two cycles, alternate the datasets into two scratch dataset systems is something that would be a simple job for a tool working on generating filenames and storage. However, manually, that implies at having at least two scripts with information about the different storage and extra work to have better utilisation of the available resources.

@@

Figure \ref{fig:stages} illustrates the stages of a workflow being processed by Cylc. Each stage has a description about what happens with the current methodologies and the increments we propose in this vision paper.

\begin{figure}[H]
  \centering
  \includegraphics[scale=1.4]{stages}
  \caption{Stages}
  \label{fig:stages}
\end{figure}

@@

  \textbf{Suite + I/O Vision}\ \ Together with the file \texttt{suite.rc}, the user will be responsible to insert the file \texttt{suite-io.rc}, as described in Figure/Section.

@@

  \textbf{Cylc + I/O Vision}\ \ Cylc will also call EWIOScheduler to potentially reprioritise the execution order of the tasks.
  EWIOScheduler will read both suite files to make suggestions, without enforcing any changings at this moment {\color{cyan}{why?}}.

  \item \textbf{Slurm}\ \ Once a task can be executed (dependencies are fulfilled), Cylc generates on the fly a Slurm script with the required metadata for Slurm, invokes the user-defined script for that task, and runs Slurm to queue the process.

  \textbf{Slurm + I/O Vision}\ \ Assuming the suggestions made by EWIOScheduler are acepted, Cylc will change the metadata of the Slurm script. Optimisation in this stage can be, for instance, defining places for a job on the same nodes as a previous job to reuse local data {\color{cyan}{NextgenIO?}} and adding to the metadata information about migrating the data.

@@

  \textbf{Task + I/O Vision}\ \ The optimisation defined in the previous stage are implemented. For instance, if migrations have to be performed before or after a job is run, \sout{{\color{cyan}{it}}} {\color{cyan}{Slurm}} will administer them.

@@

  \textbf{Script + I/O Vision}\ \ The way the filenames are generated are now slightly different. We will use a new command that invokes ESWIOSchedule again {\color{cyan}{Slurm is calling ESWIOSchedule?}} to read the suite files and generate information for ESDM about where to prioritise {\color{cyan}{prioritise ==> Is this the right word here?}} data placement according to data accessment. The filename hence encodes additional information.
  Potentially, we also can generate an \texttt{esdm.conf} file to reflect the specific hardware we have available to run the workflow and employ ESDM to perform the I/O.

@@

  \textbf{Application + I/O Vision}\ \ If the \texttt{esdm.conf} file is generated in the last stage, now the application can also use ESDM to perform the I/O. Knowing the potential targets, ESDM loads the system configuration to schedule application-local I/O. The application uses ESDM with the script filename, ESDM extracts the long-term schedule information and considers it during I/O scheduling to exploit the data locality. A data management service may run and purge or migrate data according to \sout{{\color{blue}{user-specifications}}} {\color{cyan}{user or ESDM specifications}} of the information life cycle.

@@

\begin{figure}[H]
  \centering
  \includegraphics[scale=0.6]{workflow-all}
  \caption{Workflow All.}
  \label{fig:work2}
\end{figure}

@@

\item[Input] The information inserted in the workflow.

\item[Tasks] The tasks specified by the user.

\item[Products] Any of the possible outcomes of the tasks.

For simplicity, we will call these datasets as \texttt{product-workflow.ds}. In this specific example, it is composed by the datasets \texttt{product1.ds}, \texttt{product2.ds}, and \texttt{product3.ds}.
